### 2016-04-13 sq_mobility/sq_mobility_twitter.tex
init

\documentclass[a4paper,11pt]{article}
\begin{document}
\title{Mining sequential mobility patterns from semantic Twitter user trajectories}
%\author{names}
\maketitle

\section*{Abstract}

\section*{Introduction}
Understanding detailed human mobility patterns is great importance for

\section*{Methods and Materials}

\end{document}