### 2016-05-30 NY_taxi_Flow/NY_Taxi_Flow_Capacity.tex
new content

\documentclass[a4paper,11pt]{article}
\begin{document}
%\title{A visual-analytic approach to understand New York taxi data with flow dimension and capacity}
\title{A visual-analytics approach to studying spatiotemporal taxi mobility patterns in New York}
\maketitle

\section*{\centering Abstract}

\section{Introduction}
Talk about origin and destination flow
Migration, traffic management

Hierarchical

In contrast to the network derived from

\section{Related Work}

\section{Methods and Materials}

\end{document}