### 2016-09-23 semantic_decomposition_of_twitter_user_locations/paper.tex
edittings

\documentclass[a4paper,11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{lastpage,fancyhdr,graphicx}
\graphicspath{ {/figure/} }
\usepackage{authblk}

\begin{document}
% \title{A data synthesis approach for semantic decompositions of Twitter user locations: A case study of geo-located Twitter data in Chicago}
\title{A tale of two cities: A GIS based data synthesis approach for semantic decompositions of Twitter user locations}
\author[1~\thanks{jyn@illinois.edu}]{Junjun Yin}
\affil[1]{Department of Geography and Geographic Information Science}
\affil[ ]{University of Illinois at Urbana-Champaign, IL, 61801, USA}
\renewcommand\Authands{ and }
\maketitle

\section*{Abstract}
Today's pervasive Location Based Social Media provide abundant user-generated geographic information
In this paper, we present a scalable data integration 
We analyzed the

\section{Introduction}
Modern urban environments are perceived as real-world complex systems, where the embedded urban dynamics are interrelated to various types of human activities that are taking place simultaneously~\cite{jiangyin2014}.
Throughout the process of urbanization, a fascinating research direction remains open regarding whether the spatial configurations of urban environment confine human activities or the social dynamics embedded in human activities change the functions of urban regions and therefore shape the urban structures.
Research efforts in understanding the interactions between urban environment and its citizens are of critical importance to urban planing and its applications concerning the characteristics of urban mobility, accessibility and sustainability.

what is it?

why it is important?

why is my solution?

The remainder of this paper is organized as follows. Section 2 introduces the related work 
Section 3 details the 
Detailed analysis of the
Section 5 presents the results
Section 6 concludes the paper.

\section{Related Work}

\section{Methods and Materials}
\subsection{Geo-located Twitter Data}
Geo-located Twitter data refer to

\subsection{A scalable data synthesis framework}

\section{Results}

\section{Discussions and Conclusions}
Talk about OpenStreetMap for potential large scale validation
talk about the limitation of detailed land use maps currently available.

\bibliographystyle{apalike}
\bibliography{ref}

\end{document}