### 2016-10-19 Multi-scale_Mobility_Pattern_CEUS/template-num.tex
init

% Template article for preprint document class `elsart'
% SP 2006/04/26

\documentclass{elsart}

% Use the option doublespacing or reviewcopy to obtain double line spacing
% \documentclass[doublespacing]{elsart}

% if you use PostScript figures in your article
% use the graphics package for simple commands
% \usepackage{graphics}
% or use the graphicx package for more complicated commands
% \usepackage{graphicx}
% or use the epsfig package if you prefer to use the old commands
% \usepackage{epsfig}

% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}

% The lineno packages adds line numbers. Start line numbering with
% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
% for the whole article with \linenumbers.
% \usepackage{lineno}

% \linenumbers
\begin{document}

\begin{frontmatter}

% Title, authors and addresses

% use the thanksref command within \title, \author or \address for footnotes,
% use the corauthref command within \author for corresponding author footnotes,
% use the ead command for the email address,
% and the form \ead[url] for the home page:
% \title{Title\thanksref{label1}}
% \thanks[label1]{}
% \author{Name\corauthref{cor1}\thanksref{label2}}
% \ead{email address}
% \ead[url]{home page}
% \thanks[label2]{}
% \corauth[cor1]{}
% \address{Address\thanksref{label3}}
% \thanks[label3]{}

\title{}

% use optional labels to link authors explicitly to addresses:
% \author[label1,label2]{}
% \address[label1]{}
% \address[label2]{}

\author{}

\address{}

\begin{abstract}
% Text of abstract
\end{abstract}

\begin{keyword}
% keywords here, in the form: keyword \sep keyword

% PACS codes here, in the form: \PACS code \sep code
\PACS 
\end{keyword}
\end{frontmatter}

% main text
\section{}
\label{}

% The Appendices part is started with the command \appendix,
% appendix sections are then done as normal sections
% \appendix

% \section{}
% \label{}

\begin{thebibliography}{00}
\bibitem{zheng2008understanding}
Zheng, Y., Li, Q., Chen, Y., Xie, X., Ma, W.Y., 2008
\newblock Understanding mobility based on GPS data.
\newblock  Proceedings of the 10th international conference on Ubiquitous
  computing. ACM, pp. 312--321.

\bibitem{jiang2009characterizing}
Jiang, B., Yin, J., Zhao, S., 2009
\newblock Characterizing the human mobility pattern in a large street network.
\newblock {\em Physical Review E}, {\em 80},~021136.

\bibitem{belik2011natural}
Belik, V., Geisel, T., Brockmann, D.
\newblock Natural human mobility patterns and spatial spread of infectious
  diseases.
\newblock {\em Physical Review X} {\bf 2011}, {\em 1},~011001.

\bibitem{greenwood1985human}
Greenwood, M.J.
\newblock Human migration: Theory, models, and empirical studies.
\newblock {\em Journal of regional Science} {\bf 1985}, {\em 25},~521--544.

\bibitem{brockmann2006scaling}
Brockmann, D., Hufnagel, L., Geisel, T.
\newblock The scaling laws of human travel.
\newblock {\em Nature} {\bf 2006}, {\em 439},~462--465.

\bibitem{gonzalez2008understanding}
Gonzalez, M.C., Hidalgo, C.A., Barabasi, A.L.
\newblock Understanding individual human mobility patterns.
\newblock {\em Nature} {\bf 2008}, {\em 453},~779--782.

\bibitem{Jurdak2015}
Jurdak, R., Zhao, K., Liu, J., AbouJaoude, M., Cameron, M., Newth, D.
\newblock Understanding Human Mobility from Twitter.
\newblock {\em PLoS ONE} {\bf 2015}, {\em 10},~e0131469.

\bibitem{rhee2011levy}
Rhee, I., Shin, M., Hong, S., Lee, K., Kim, S.J., Chong, S.
\newblock On the levy-walk nature of human mobility.
\newblock {\em IEEE/ACM transactions on networking (TON)} {\bf 2011}, {\em
  19},~630--643.

\bibitem{sevtsuk2010does}
Sevtsuk, A., Ratti, C.
\newblock Does urban mobility have a daily routine? Learning from the aggregate
  data of mobile networks.
\newblock {\em Journal of Urban Technology} {\bf 2010}, {\em 17},~41--60.

\bibitem{kung2014exploring}
Kung, K.S., Greco, K., Sobolevsky, S., Ratti, C.
\newblock Exploring universal patterns in human home-work commuting from mobile
  phone data.
\newblock {\em PloS one} {\bf 2014}, {\em 9},~e96180.

\bibitem{thatcher2014living}
Thatcher, J.
\newblock Living on fumes: Digital footprints, data fumes, and the limitations
  of spatial big data.
\newblock {\em International Journal of Communication} {\bf 2014}, {\em
  8},~1765--1783.

\bibitem{hawelka2014geo}
Hawelka, B., Sitko, I., Beinat, E., Sobolevsky, S., Kazakopoulos, P., Ratti, C.
\newblock Geo-located Twitter as proxy for global mobility patterns.
\newblock {\em Cartography and Geographic Information Science} {\bf 2014}, {\em
  41},~260--271.

\bibitem{giannotti2008mobility}
Giannotti, F., Pedreschi, D.
\newblock {\em Mobility, data mining and privacy: Geographic knowledge
  discovery}, Springer Science \& Business Media,  2008.

\bibitem{crampton2014collect}
Crampton, J.W.
\newblock Collect it all: national security, Big Data and governance.
\newblock {\em GeoJournal} {\bf 2014}, pp. 1--13.

\bibitem{wu2014intra}
Wu, L., Zhi, Y., Sui, Z., Liu, Y.
\newblock Intra-urban human mobility and activity transition: Evidence from
  social media check-in data.
\newblock {\em PloS one} {\bf 2014}, {\em 9},~e97010.

\bibitem{hasan2013understanding}
Hasan, S., Zhan, X., Ukkusuri, S.V.
\newblock Understanding urban human activity and mobility patterns using
  large-scale location-based data from online social media.
\newblock  Proceedings of the 2nd ACM SIGKDD international workshop on urban
  computing. ACM,  2013, p.~6.

\bibitem{cho2011friendship}
Cho, E., Myers, S.A., Leskovec, J.
\newblock Friendship and mobility: user movement in location-based social
  networks.
\newblock  Proceedings of the 17th ACM SIGKDD international conference on
  Knowledge discovery and data mining. ACM,  2011, pp. 1082--1090.

\bibitem{noulas2012tale}
Noulas, A., Scellato, S., Lambiotte, R., Pontil, M., Mascolo, C.
\newblock A tale of many cities: universal patterns in human urban mobility.
\newblock {\em PloS one} {\bf 2012}, {\em 7},~e37027.

\bibitem{balcan2009multiscale}
Balcan, D., Colizza, V., Gon{\c{c}}alves, B., Hu, H., Ramasco, J.J.,
  Vespignani, A.
\newblock Multiscale mobility networks and the spatial spreading of infectious
  diseases.
\newblock {\em Proceedings of the National Academy of Sciences} {\bf 2009},
  {\em 106},~21484--21489.

\bibitem{tamerius2011global}
Tamerius, J., Nelson, M.I., Zhou, S.Z., Viboud, C., Miller, M.A., Alonso, W.J.
\newblock Global influenza seasonality: reconciling patterns across temperate
  and tropical regions.
\newblock {\em Environmental health perspectives} {\bf 2011}, {\em 119},~439.

\bibitem{tsou2015}
Tsou, M.H.
\newblock Research challenges and opportunities in mapping social media and Big
  Data.
\newblock {\em Cartography and Geographic Information Science} {\bf 2015}, {\em
  42},~70--74.

\bibitem{zheng2010geolife}
Zheng, Y., Xie, X., Ma, W.Y.
\newblock GeoLife: A Collaborative Social Networking Service among User,
  Location and Trajectory. {\bf 2010}.

\bibitem{becker2013human}
Becker, R., C{\'a}ceres, R., Hanson, K., Isaacman, S., Loh, J.M., Martonosi,
  M., Rowland, J., Urbanek, S., Varshavsky, A., Volinsky, C.
\newblock Human mobility characterization from cellular network data.
\newblock {\em Communications of the ACM} {\bf 2013}, {\em 56},~74--82.

\bibitem{sobolevsky2013delineating}
Sobolevsky, S., Szell, M., Campari, R., Couronn{\'e}, T., Smoreda, Z., Ratti,
  C.
\newblock Delineating geographical regions with networks of human interactions
  in an extensive set of countries.
\newblock {\em PLoS One} {\bf 2013}, {\em 8},~e81707.

\bibitem{twitterAPI}
Twitter.
\newblock Twitter streaming API.
\newblock {\em available from: https://dev.twitter.com/streaming/overview} {\bf
  2016}.

\bibitem{cranshaw2012livehoods}
Cranshaw, J., Schwartz, R., Hong, J.I., Sadeh, N.M.
\newblock The Livehoods Project: Utilizing Social Media to Understand the
  Dynamics of a City.
\newblock  ICWSM,  2012.

\bibitem{mitchell2013geography}
Mitchell, L., Frank, M.R., Harris, K.D., Dodds, P.S., Danforth, C.M.
\newblock The geography of happiness: Connecting twitter sentiment and
  expression, demographics, and objective characteristics of place {\bf 2013}.

\bibitem{longley2015geotemporal}
Longley, P.A., Adnan, M., Lansley, G., others.
\newblock The geotemporal demographics of Twitter usage.
\newblock {\em Environment and Planning A} {\bf 2015}, {\em 47},~465--484.

\bibitem{hagerstrand1985time}
H{\"a}gerstrand, T., others.
\newblock Time-geography: focus on the corporeality of man, society, and
  environment.
\newblock {\em The science and praxis of complexity} {\bf 1985}, pp. 193--216.

\bibitem{kwan2004geovisualization}
Kwan, M.P., Lee, J.
\newblock Geovisualization of human activity patterns using 3D GIS: a
  time-geographic approach.
\newblock {\em Spatially integrated social science} {\bf 2004}, {\em 27}.

\bibitem{andrienko2007designing}
Andrienko, N., Andrienko, G.
\newblock Designing visual analytics methods for massive collections of
  movement data.
\newblock {\em Cartographica: The International Journal for Geographic
  Information and Geovisualization} {\bf 2007}, {\em 42},~117--138.

\bibitem{maceachren2001research}
MacEachren, A.M., Kraak, M.J.
\newblock Research challenges in geovisualization.
\newblock {\em Cartography and Geographic Information Science} {\bf 2001}, {\em
  28},~3--12.

\bibitem{maceachren2004maps}
MacEachren, A.M.
\newblock {\em How maps work: representation, visualization, and design},
  Guilford Press,  2004.

\bibitem{andrienko2007visual}
Andrienko, G., Andrienko, N., Wrobel, S.
\newblock Visual analytics tools for analysis of movement data.
\newblock {\em ACM SIGKDD Explorations Newsletter} {\bf 2007}, {\em 9},~38--46.

\bibitem{black2012twitter}
Black, A., Mascaro, C., Gallagher, M., Goggins, S.P.
\newblock Twitter zombie: Architecture for capturing, socially transforming and
  analyzing the Twittersphere.
\newblock  Proceedings of the 17th ACM international conference on Supporting
  group work. ACM,  2012, pp. 229--238.

\bibitem{shvachko2010hadoop}
Shvachko, K., Kuang, H., Radia, S., Chansler, R.
\newblock The hadoop distributed file system.
\newblock  Mass Storage Systems and Technologies (MSST), 2010 IEEE 26th
  Symposium on. IEEE,  2010, pp. 1--10.

\bibitem{dean2008mapreduce}
Dean, J., Ghemawat, S.
\newblock MapReduce: simplified data processing on large clusters.
\newblock {\em Communications of the ACM} {\bf 2008}, {\em 51},~107--113.

\bibitem{gao2012exploring}
Gao, H., Tang, J., Liu, H.
\newblock Exploring Social-Historical Ties on Location-Based Social Networks.
\newblock  ICWSM,  2012.

\bibitem{buttenfield1991map}
Buttenfield, B.P., McMaster, R.B.
\newblock {\em Map Generalization: Making rules for knowledge representation},
  Longman Scientific \& Technical New York,  1991.

\bibitem{samet1984quadtree}
Samet, H.
\newblock The quadtree and related hierarchical data structures.
\newblock {\em ACM Computing Surveys (CSUR)} {\bf 1984}, {\em 16},~187--260.

\bibitem{clauset2009power}
Clauset, A., Shalizi, C.R., Newman, M.E.
\newblock Power-law distributions in empirical data.
\newblock {\em SIAM review} {\bf 2009}, {\em 51},~661--703.

\bibitem{reynolds2012truncated}
Reynolds, A.
\newblock Truncated L{\'e}vy walks are expected beyond the scale of data
  collection when correlated random walks embody observed movement patterns.
\newblock {\em Journal of The Royal Society Interface} {\bf 2012}, {\em
  9},~528--534.

\bibitem{zhao2015explaining}
Zhao, K., Musolesi, M., Hui, P., Rao, W., Tarkoma, S.
\newblock Explaining the power-law distribution of human mobility through
  transportation modality decomposition.
\newblock {\em Scientific reports} {\bf 2015}, {\em 5}.

\end{thebibliography}

\end{document}