### 2016-12-27 semantic_decomposition_of_twitter_user_locations/paper.tex
test

\documentclass[a4paper, 11pt]{article}
\usepackage{authblk}
\usepackage[round, sort, numbers, authoryear]{natbib}
\usepackage{regexpatch}
\usepackage{graphicx}
\graphicspath{ {/figures/} }
\usepackage[section]{placeins}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\usepackage[]{algorithm2e}
\usepackage{caption}
\usepackage{subcaption}
\date{}

\usepackage{lineno}

\begin{document}
% \title{A data synthesis approach for semantic decompositions of Twitter user locations: A case study of geo-located Twitter data in Chicago}
\title{A tale of two cities: A GIS based data synthesis approach to understanding semantic decompositions of Twitter user locations}
% \author~\thanks{yinjunjun@gmail.com]{Junjun Yin}
\author[1,~\thanks{Corresponding author: jyin@psu.edu}]{Junjun Yin}
\affil[1]{Population Research Institute}
\affil[ ]{The Pennsylvania State University, State College, 16801, PA, USA}
\renewcommand\Authands{ and }
\maketitle

\section*{Abstract}
Today's pervasive Location Based Social Media provide abundant user-generated geographic information
In this paper, we present a geographical information system (GIS) based data synthesis approche to understanding semantic decompositions of Twitter user locations.
Specifically, we implemented a
This approach 
We analyzed the

\section{Introduction}

xx~\citep{yin2016exploring}

With the advancements in mobile technology, the location information 
what is it?

why it is important?

why is my solution?

The remainder of this paper is organized as follows. Section 2 introduces the related work 
Section 3 details the 
Detailed analysis of the
Section 5 presents the results
Section 6 concludes the paper.

\section{Related Work}

\section{Methods and Materials}

\subsection{Geo-located Twitter Data}
Geo-located Twitter data refer to

\subsection{A scalable data synthesis framework}

\section{Results}

\section{Discussions and Conclusions}
Talk about OpenStreetMap for potential large scale validation
talk about the limitation of detailed land use maps currently available.

\bibliographystyle{apalike}
\bibliography{ref}
\end{document}